A highly-available move operation for replicated trees

### 2021-08-18 paper/review-response2.tex

\usepackage{xr}
\externaldocument{move-op}

\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}
\newcommand{\authorcomment}[1]{\begin{quote}\textbf{Author comment:} #1\end{quote}}

\begin{document}
\title{Summary of changes to ``A highly-available move operation for replicated trees''}
\author{Martin Kleppmann, Dominic P.\ Mulligan, Victor B.F.\ Gomes, Alastair R.\ Beresford}
\date{}
\maketitle

We thank all reviewers for this second round of reviews and are happy to note that reviewers 2 and 3 both would like to see the paper accepted as-is with no further changes.
In this revision we have therefore only made minor changes to address comments by reviewer 1, and added some minor clarifications.

@@

\section{Reviewer 1}

\begin{spverbatim}
Recommendation: Author Should Prepare A Major Revision For A Second Review

Comments:
What is the difference between the proposed algorithm and the data replication algorithms in the cloud? Please state their challenges in the introduction by mentioning references.
A review of data replication based on meta-heuristics approach in cloud computing and data grid. doi.org/10.1007/s00500-020-04802-1.
\end{spverbatim}
\authorcomment{The cited paper is about using optimisation algorithms to decide what to replicate, how many replicas to create, where to locate them, when to access them, and other similar questions.
Whilst we agree that those questions are indeed important when deciding how to deploy real-world systems we think that they are unrelated to our work, which is a largely theoretical work introducing (and thereafter proving correct) a novel distributed algorithm.
In particular, our paper introduces techniques that ensure that all replicas converge to a consistent ``view'' of a distributed tree data structure.
Our work can be deployed in any system that requires the use of a replicated tree, regardless of whether it is cloud-based, client-server, peer-to-peer, or uses any other topology---from our point-of-view we are completely ambivalent about such deployment questions as they exist at a level of abstraction \emph{below} that in which we are working.
Simply put: there is no need for us to distinguish between cloud systems and other types of replicated system because from the point of view of our algorithm they are the same.}
\begin{spverbatim}
A brief survey on replica consistency in cloud environments. doi.org/10.1186/s13174-020-0122-y
\end{spverbatim}
\authorcomment{As we explain in {\S}1, {\S}4.2, and {\S}6, our algorithm provides a consistency model called ``strong eventual consistency'' [Shapiro et al., 2011].
The survey referenced above unfortunately overlooks this important consistency model.
Strong eventual consistency is a strengthening of eventual consistency, and it is often combined with causal consistency, and it is one of the strongest consistency models that is always provides ``availability'' in the sense of the CAP theorem.
We believe our paper already adequately explains the consistency model, and the prior works [2,3,4] referenced in our paper explain it in further detail.}
\begin{spverbatim}
Explain the time complexity of the proposed method.
\end{spverbatim}
\authorcomment{At the request of the reviewer we have added a discussion of the time complexity of our algorithm to the beginning of {\S}5.1.}
\begin{spverbatim}
Unfortunately, there are no "quantitative" performance comparisons among related strategies. It’s also important if the author can summarize the advantages, and disadvantages of each strategy according to the performance metrics.
\end{spverbatim}
\authorcomment{Our evaluation ({\S}5) performs a quantitative performance comparison between two implementations of our algorithm and state machine replication.
We explain the advantages and disadvantgaes of these approaches in detail in this section.
In \S~5.2 we also compare our algorithm to a locking-based approach, and show that it would have three orders of magnitude lower throughput than our algorithm in our test setup (5.7 versus 5,700 ops/sec).

We therefore believe that our paper already contains the quantitative comparison that the reviewer is asking for.
If there is a particular performance comparison that is missing in the reviewer's opinion, we kindly request that they specify precisely what should be added.}
\begin{spverbatim}

### 2021-08-18 paper/review-response2.tex

We thank all reviewers for this second round of reviews and are happy to note that reviewers 2 and 3 both would like to see the paper accepted as-is with no further changes.
In this revision we have therefore only made minor changes to address comments by reviewer 1, and added some minor clarifications.

\section{Reviewer 1}

\begin{spverbatim}
Recommendation: Author Should Prepare A Major Revision For A Second Review

Comments:
What is the difference between the proposed algorithm and the data replication algorithms in the cloud? Please state their challenges in the introduction by mentioning references.
A review of data replication based on meta-heuristics approach in cloud computing and data grid. doi.org/10.1007/s00500-020-04802-1.
\end{spverbatim}
\authorcomment{The cited paper is about using optimisation algorithms to decide what to replicate, how many replicas to create, where to locate them, when to access them, and other similar questions.
Whilst we agree that those questions are indeed important when deciding how to deploy real-world systems we think that they are unrelated to our work, which is a largely theoretical work introducing (and thereafter proving correct) a novel distributed algorithm.
In particular, our paper introduces techniques that ensure that all replicas converge to a consistent ``view'' of a distributed tree data structure.
Our work can be deployed in any system that requires the use of a replicated tree, regardless of whether it is cloud-based, client-server, peer-to-peer, or uses any other topology---from our point-of-view we are completely ambivalent about such deployment questions as they exist at a level of abstraction \emph{below} that in which we are working.
Note that, whilst we provide an empirical assessment of our algorithm, this assessment is merely intended to demonstrate that our algorithm is not \emph{ipso facto} unworkably inefficient.

Simply put: there is no need for us to distinguish between cloud systems and other types of replicated system because from the point of view of our algorithm they are the same.}
\begin{spverbatim}
A brief survey on replica consistency in cloud environments. doi.org/10.1186/s13174-020-0122-y
\end{spverbatim}
\authorcomment{As we explain in {\S}1, {\S}4.2, and {\S}6, our algorithm provides a consistency model called ``strong eventual consistency'' [Shapiro et al., 2011].
The survey referenced above unfortunately overlooks this important consistency model.
Strong eventual consistency is a strengthening of eventual consistency, is often combined with causal consistency, and it is one of the strongest consistency models that is always provides ``availability'' in the sense of the CAP theorem.
We believe our paper already adequately explains the consistency model, and the prior works [2,3,4] referenced in our paper explain it in further detail.}
\begin{spverbatim}
Explain the time complexity of the proposed method.
\end{spverbatim}
\authorcomment{At the request of the reviewer we have added a discussion of the time complexity of our algorithm to the beginning of {\S}5.1.}
\begin{spverbatim}
Unfortunately, there are no "quantitative" performance comparisons among related strategies. It’s also important if the author can summarize the advantages, and disadvantages of each strategy according to the performance metrics.
\end{spverbatim}
\authorcomment{Our evaluation ({\S}5) performs a quantitative performance comparison between two implementations of our algorithm and state machine replication.
We explain the advantages and disadvantgaes of these approaches in detail in this section.
In \S~5.2 we also compare our algorithm to a locking-based approach, and show that it would have three orders of magnitude lower throughput than our algorithm in our test setup (5.7 versus 5,700 ops/sec).

### 2021-10-04 paper/move-op.tex
Terminology tweak

With our algorithm, the worst-case cost of applying a move operation is $O(hd)$, where $h$ is the number of operations in the history that need to be undone and redone, and $d$ is the depth of the tree (the number of parent relationships that need to be traversed to check whether the operation would introduce a cycle).
To demonstrate that this cost is acceptable in practice, we evaluated the performance of two implementations of our algorithm.
The first implementation is Scala code extracted from our formally verified HOL definitions using Isabelle/HOL's code generation mechanism.
The second implementation is handwritten Scala code that we believe to be functionally equivalent, but which we have not formally verified.
By comparing these implementations we can distinguish between inefficiencies introduced by code generation and costs that are inherent to our algorithm.

>>>>

With our algorithm, the worst-case cost of applying a move operation is $O(nd)$, where $n$ is the number of operations in the log that need to be undone and redone, and $d$ is the depth of the tree (the number of parent relationships that need to be traversed to check whether the operation would introduce a cycle).
To demonstrate that this cost is acceptable in practice, we evaluated the performance of two implementations of our algorithm.
The first implementation is Scala code extracted from our formally verified HOL definitions using Isabelle/HOL's code generation mechanism.
The second implementation is handwritten Scala code that we believe to be functionally equivalent, but which we have not formally verified.
By comparing these implementations we can distinguish between inefficiencies introduced by code generation and costs that are inherent to our algorithm.
